---
title: "HFMM Model Fitting"
author: "Qian Li"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this writeup, we compare three modelling strategies:

- Strongly separable model
- Vectorized Latent Factor model
- Sandwich Latent Factor model

## Simulation Setups
We simulate the band powers over 6 regions, that has a predetermined AR(1) covariance structure.
```{r}
# ------- Simulation parameters ----------
nsub <-       10          # Number of subjects per group
ind.smooth <- 2           # Smoothness of the GP, ++ for smoother deviation
ind.var <-    .8          # Variance of the GP (SNR related)
noise.var <-  1           # Variance of the error term (SNR related)
mis.p <-      .4          # Missing probability for each time point
AR.rho <-     .8          # Rho for the AR cov in simulation
num.knots <-  10          # Number of knots
# ---------------------------------------------------------------------------+
# To knit: rmarkdown::render('vignettes/fit-test.Rmd')
```

```{r, include = FALSE, warning=FALSE}
### Loading Required packages
require(Matrix); require(MASS); require(reshape2); require(ggplot2)
require(lattice); library(HFMM); require(pander)
### Curves and time
time <- seq(0,30, by = .5)
curves <- list(
  curve0 = function(x) {rep(1, length(x))},
  curve1 = function(x) {5 * sin(x * pi / 60)},
  curve2 = function(x) {2 - 3 * cos((x - 25) * pi / 15)},
  curve3 = function(x) {5 - 0.2 * x},
  curve4 = function(x) {-1 + (20 - x) ^ 2 / 100}
)
## Cov: AR(1)
cov <- abs(outer(1:6,1:6,"-")); sp_cov <- AR.rho ^ cov; chol_sp <- chol(sp_cov)
## Group means
gcurves <- list()
gcurves[[1]] <- lapply(curves[c(1,1,1,1,1,1)], function(f) f(time))
gcurves[[2]] <- lapply(curves[c(2,2,2,3,3,4)], function(f) f(time))
gcurves[[3]] <- lapply(curves[c(3,4,5,4,5,5)], function(f) f(time))
gcurves <- lapply(gcurves, function(x) matrix(unlist(x), ncol = 6, byrow = F))
## Gaussian Process covariance
# GP.cov <- outer(time, time, "-"); GP.cov <- exp(-1/2 * (abs(GP.cov)/ind.smooth)^2)
GP.cov <- HFMM::calcSigma(time, time, l = ind.smooth)
## Missing labels (const)
p <- rep(mis.p,nsub*3*length(time)); p <- (runif(length(p)) >= p)
p <- split(p, rep(1:(nsub*3),rep(length(time),nsub*3)))
## Covariates and simulation settings
sim_time <- unlist(lapply(p, function(x) time[x]))
sim_nsegs <- unname(unlist(lapply(p, sum)))
sim_group <- rep(c(1,2,3), rep(nsub,3))
sim_knots <- default.knots(sim_time, num.knots = num.knots)
```

### Simulated Group Means for 6 Regions.
```{r, echo = FALSE, fig.height=4, fig.width=6}
par(mfrow = c(2,3), mar = c(1,1,3,1))
for(i in 1:6){
  matplot(time,matrix(unlist(lapply(gcurves, function(x) x[,i])), ncol = 3), type='l', lwd = 2, xlab = '', main = paste0("Region ",i), col = c('blue','black','red'))
}   
```

Within each region, the group means are selected from a pool of predetermined curves:

$$
\beta_0(t) = 1; \quad \beta_1(t) = 5\cdot\sin(t*\pi/60); \quad \beta_2(t) =  2-3\cdot \cos((t-25)*\pi/15)\\
\beta_3(t) = 5-0.2t;\quad  \beta_4(t) = -1+(20-t)^2/100,
$$
therefore, 

* $\text{Region 1}:\{\beta_0(t), \beta_1(t), \beta_2(t)\}$
* $\text{Region 2}:\{\beta_0(t), \beta_1(t), \beta_3(t)\}$
* $\text{Region 3}:\{\beta_0(t), \beta_1(t), \beta_4(t)\}$
* $\text{Region 4}:\{\beta_0(t), \beta_2(t), \beta_3(t)\}$
* $\text{Region 5}:\{\beta_0(t), \beta_2(t), \beta_4(t)\}$
* $\text{Region 6}:\{\beta_0(t), \beta_3(t), \beta_4(t)\}$

```{r, echo = FALSE, fig.width=8, fig.height=8}
complete_data <- list()
for(g in 1:3){
  for(sub in 1:nsub){
    index <- sub + (g-1) * nsub
    bij <- mvrnorm(6,rep(0,length(time)), Sigma = GP.cov*ind.var)
    complete_data[[index]] <- gcurves[[g]] + t(bij) %*% chol_sp + 
      t(mvrnorm(6,rep(0,length(time)), Sigma = noise.var*diag(length(time))))
  }
}
for(i in 1:length(complete_data)) complete_data[[i]] <- complete_data[[i]][p[[i]],]
sim_data <- do.call(rbind, lapply(complete_data, as.matrix))

## Model 1 fit
# X <- matrix(0, nrow=length(sim_group), ncol=3)
# X[sim_group==1,1] <- 1
# X[sim_group==2,2] <- 1
# X[sim_group==3,3] <- 1
# mc1 <- list(burnin = 1000, nsim = 5000, thin = 2)
# sim_fit <- hfm(sim_data, sim_time, X, nsegs=sim_nsegs, knots=sim_knots, mcmc=mc1)
## Model 2 fit
X <- matrix(0, nrow=length(sim_group), ncol=3)
X[sim_group==1,1] <- 1
X[sim_group==2,2] <- 1
X[sim_group==3,3] <- 1
mc1 <- list(burnin = 0, nsim = 1000, thin = 1)
sim_fit <- vfm(sim_data, sim_time, X, 10, nsegs=sim_nsegs, knots=sim_knots, mcmc=mc1)
## Model 3 fit
```

## Fitting Visualization, sub-region curves
```{r, echo = FALSE, fig.width=9, fig.height=9}
par(mfrow=c(3,3))
# tt <- sort(unique(as.vector(sim_time)))/30;
sampID <- c();  for(g in 1:3) sampID <- c(sampID, sample(nsub,1)+(g-1)*nsub)
for(i in sampID){
  for(j in c(1,3,6)){
    plot(time[p[[i]]], complete_data[[i]][,j], cex = 0.8, xlab = 'Time', ylab = 'Signal')
    # plot(sim_fit$t[i,],sim_fit$y[i,j,], cex=0.8, xlab='Normalized Time', ylab='Signal')
    title(paste(paste("Subject", i, "-"), paste("Region", j)))
    # lines(tt,sim_fit$fit[i,j,], col=4)
  }
}
```

## Fitting Visualization, group-region curves
